Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-10-29 paper/results.tex
Adding tables of results; need to add the numbers and some text still.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@647 3fc257bb-8550-0410-81b3-a5c103da83f7

Figure~\ref{f:pr} shows the execution time per PageRank iteration for
R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  Converging the PageRank
iterations to tolerance $0.002$ requires five or six iterations.
Several R-MAT matrices of each size were generated for the
experiments; the average time over the matrices is reported here.  The
Trilinos implementations show near-perfect strong scaling for RMAT-20
and RMAT-24.  The MR-MPI implementations also demonstrate good strong
scaling.  However, MR-MPI's execution time is at least an order of
magnitude greater than the Trilinos implementation.  This result is
due to two factors: ({\it i}) a higher volume of communication in the
MapReduce implementation (where all edges are communicated in each
iteration), compared to the Trilinos implementation (where only
vertex-based data is communicated), and ({\it ii}) out-of-core
operations in MR-MPI were performed because of the page-size
restriction, while all Trilinos operations were performed in-core.
The benefit of MR-MPI's out-of-core implementation is seen, however,
with the RMAT-28 data set, which could be solved on smaller processor
sets than the Trilinos implementation.  For these experiments, the
Trilinos implementation required 64 processors for the RMAT-28 data
set, since it will not operate on matrices that exceed the aggregate
memory of the processors.

@@

>>>>

Figure~\ref{f:pr} shows the execution time per PageRank iteration for
R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  Converging the PageRank
iterations to tolerance $0.002$ requires five or six iterations.
Several R-MAT matrices of each size were generated for the
experiments; the average time over the matrices is reported here.  The
Trilinos implementations show near-perfect strong scaling for RMAT-20
and RMAT-24.  The MR-MPI implementations also demonstrate good strong
scaling.  However, MR-MPI's execution time is at least an order of
magnitude greater than the Trilinos implementation.  
The benefit of MR-MPI's out-of-core implementation
is seen, however, with the RMAT-28 data set, which could be solved on
smaller processor sets than the Trilinos implementation.  For these
experiments, the Trilinos implementation required 64 processors for
the RMAT-28 data set, since it will not operate on matrices that
exceed the aggregate memory of the processors.

@@

A more detailed analysis of the performance of the MR-MPI PageRank is shown
below.  We perform the analysis using RMAT-20, which is small enough to
fit in memory and thus does not need out-of-core operations, and RMAT-28,
which requires out-of-core operation in MR-MPI.  Table~\ref{table:prdetail}
shows the amount of time spent in each step of Figure~\ref{fig:pr2} for 
the MR-MPI implementation and the Trilinos implementation.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline 
 & \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
 & MR-MPI & Trilinos & MR-MPI & Trilinos  \\
\hline
Step 1 (Compute adjustment) & & & &  \\
Step 2 (Multiply $A^T x$) & & & &  \\
Step 3 (Apply adjustment) & & & &  \\
Step 4 (Scale $y$)  & & & & \\
Step 5 (Compute residual)  & & & & \\
\hline
\end{tabular}
\caption{Detailed execution times in each stage of the PageRank algorithm
described in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
\label{table:prdetail}
\end{center}
\end{table}

In Tables~\ref{table:prdetail2},
we break down the operation times further for the most expensive 
step of the PageRank computation, Step 2.
In the table, we divide the collate operation into its
two parts:  an aggregate that involves parallel communication and a convert
that gathers local key-value pairs with matching keys into a key-multivalue
pair.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
 & RMAT-20 & RMAT-28 \\
\hline
Copy $x$ to $y$ & &  \\
Add $A$ to $y$ & &  \\
Convert $y$ & &  \\
Reduce $y$ & & \\
Aggregate $y$ (first part of collate) & & \\
Convert $y$ (second part of collate) & & \\
Reduce $y$ & & \\
\hline
\end{tabular}
\caption{Detailed execution times of MR-MPI in Step 2 of 
Figure~\ref{fig:pr2}.}
\label{table:prdetial2}
\end{center}
\end{table}

Similarly, in Table~\ref{table:prdetail5}, we show the execution times 
required for each operation in Step 5 of Figure~\ref{fig:pr2}, the 
residual computation.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
 & RMAT-20 & RMAT-28 \\
\hline
Add $y$ to $x$ & &  \\
Convert $x$ & &  \\
Reduce $x$ & & \\
MPI\_Allreduce & & \\
\hline
\end{tabular}
\caption{Detailed execution times of MR-MPI in Step 5 of 
Figure~\ref{fig:pr2}.}
\label{table:prdetial2}
\end{center}
\end{table}
%This result is
%due to two factors: ({\it i}) a higher volume of communication in the
%MapReduce implementation (which comunicates one datum per graph edge
%in each iteration), compared to the Trilinos implementation which only
%communicates a volume of data proportional to the number of vertices,
%and ({\it ii}) out-of-core operations in MR-MPI were performed because
%of the page-size restriction, while all Trilinos operations were
%performed in-core.  
%
